Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-04-15 papers/kmeans/response_to_referee.txt

    This paper presents new element abundance results using The
    Cannon, a novel, powerful method for spectroscopic abundance
    measurements. The results presented in this paper are new and
    very much relevant to the many large scale observing projects
    that are currently in operation.

    I highlight some comments for the authors to address:

    Abstract: The last sentence is not true, as the authors note
    later in the text the first blind search purely in abundance
    space was carried out by Mitschang et al 2014.

        We have referenced Mitschang et al. (2014) for their work in
        chemical tagging, but we believe the last sentence of the
        abstract is correct. To clarify: Mitschang et al. (2014) did
        perform a blind search in chemical abundances to identify
        substructure, but one of the key conclusions of their work
        was that ".. these groups may not represent distinct 
        formation sites, e.g. as in dissolved open clusters, as 
        previously thought". While the Mitschang et al. (2014) work
        showed their chemically-identified structures had similar 
        (U, V, W) space velocities (e.g., their Figure 2), those 
        velocities are not dissimilar from the background velocity
        distribution. In contrast, our work is the first blind
        search in chemical abundances that has identified structures
        that are *also* structures in phase-space: specifically
        on-sky position and radial velocities.

        Nevertheless, to avoid unwarranted claims, we have qualified
        this statement from:

        "This is the first-ever project to identify phase-space 
        structures by blind search purely in abundance space;
        the prospects for future data sets are very good."

        to:

        "This is the first-ever project to identify phase-space 
        structures at survey-scale by blind search purely in abundance space;
        the prospects for future data sets are very good."

        since the Mitschang et al. (2014) work had a sample size of
        714 stars.

@@

    Introduction: The first paragraph is missing a reference or two,
    to support claim that most stars end up in different parts of the
    Galaxy. Or this is something that is assumed and yet to be
    tested?

        The text claims that stars end up in different parts of
        phase-space in the Galaxy: that is to say that any common
        kinematic (or dynamic: actions, angles) signature between
        stars born in the same molecular gas cloud at the same time
        will eventually disperse, presumably throughout the disk.
        This line of reasoning follows from expectations that most
        stars form in clusters, but the disk is full of stars that
        do not appear to be kinematically bound to a cluster. For
        these reasons we would claim that the opening paragraph is
        more of an introduction to the expectations from star
        formation, or rather introducing the line of reasoning for
        why elemental abundances can be instructive for reconstructing
        the evolution of the Galaxy.

@@

    There are some missing references to the original demonstrations
    of chemical tagging of unbound groups, namely the Wolf 630 group
    and HR 1614 group. Those were key to building up the case for
    this "tagging" research.

        We have added the following references: De Silva et al. 
        (2007), Eggen (1970), and Bubar & King (2010).

@@

HOGG

    Section 2:

    A reference for the famously demonstrated correlations in
    globular clusters is needed.

        We have added references to: Norris & Da Costa (1995), which
        (to our knowledge) was the first to unambiguously show anti-
        correlations in Na, Al, and O, and to Carretta et al. (2009),
        which provides a more recent and comprehensive view of our
        current knowledge of elemental abundance correlations in
        gloublar clusters.

@@

    Section 3:

    The text doesn't clarify why only k-means was chosen to do the
    'tagging' given the issues discussed, was it simply easy and
    fast? Fine as the focus of the paper is the precision of the
    abundance results with Cannon, not the fidelity of group finding.

        We have added text to clarify why k-means was chosen.
        Specifically we have updated the paragraph:

        "Nothing about this algorithm or our choices are particularly
        tuned or optimized; this is in no sense the
        algorithm or the method that reveals the best structures.
        Indeed, it is a feature of this work that even the simplest, most
        generic clustering algorithm returns interesting
        structures (as shown below). The prospects for Milky Way science only 
        improve as the clustering algorithm improves."

        To now read as:

        "Nothing about this algorithm or our choices are particularly
        tuned or optimized; this is in no sense the
        algorithm or the method that reveals the best structures.
        We chose k-means as a simple and tractable approach to identifying
        clusters in high dimensional label space. In practice more complex
        clustering algorithms may perform better than k-means, or even
        tuning k-means by heuristically setting $K$ is likely to improve
        upon the results here. This \documentname\ serves as a proof of concept.
        Indeed, it is a feature of this work that even the simplest, most
        generic clustering algorithm returns interesting
        structures (as shown below). The prospects for Milky Way science only 
        improve as the clustering algorithm improves."

### 2016-04-16 papers/kmeans/response_to_referee.txt

    I highlight some comments for the authors to address:

    Abstract: The last sentence is not true, as the authors note
    later in the text the first blind search purely in abundance
    space was carried out by Mitschang et al 2014.

We have referenced Mitschang et al. (2014) for their work in chemical
tagging, but we believe the last sentence of the abstract is
correct. To clarify: Mitschang et al. (2014) did perform a blind
search in chemical abundances to identify substructure, but one of the
key conclusions of their work was that ".. these groups may not
represent distinct formation sites, e.g. as in dissolved open
clusters, as previously thought". While the Mitschang et al. (2014)
work showed their chemically-identified structures had similar (U, V,
W) space velocities (e.g., their Figure 2), those velocities are not
dissimilar from the background velocity distribution. In contrast, our
work is the first blind search in chemical abundances that has
identified structures that are *also* structures in phase-space:
specifically on-sky position and radial velocities.

Nevertheless, to avoid unwarranted claims, we have qualified this
statement in the abstract.

@@

    Introduction: The first paragraph is missing a reference or two,
    to support claim that most stars end up in different parts of the
    Galaxy. Or this is something that is assumed and yet to be
    tested?

The text claims that stars end up in different parts of phase-space in
the Galaxy: that is to say that any common kinematic (or dynamic:
actions, angles) signature between stars born in the same molecular
gas cloud at the same time will eventually disperse, presumably
throughout the disk.  This line of reasoning follows from expectations
that most stars form in clusters, but the disk is full of stars that
do not appear to be kinematically bound to a cluster. For these
reasons we would claim that the opening paragraph is more of an
introduction to the expectations from star formation, or rather
introducing the line of reasoning for why elemental abundances can be
instructive for reconstructing the evolution of the Galaxy.

    There are some missing references to the original demonstrations
    of chemical tagging of unbound groups, namely the Wolf 630 group
    and HR 1614 group. Those were key to building up the case for
    this "tagging" research.

We have added the following references: De Silva et al (2007), Eggen
(1970), and Bubar & King (2010).

### 2016-04-16 papers/kmeans/kmeans.tex
Put in 12pt manuscript format
tweaks to ARC's work

\documentclass[12pt, letterpaper, preprint]{aastex}

@@

Chemical tagging holds the promise of revealing not just the
star-formation history of the Galaxy, but also the accretion history
(as things that fall in are expected to be chemically distinct from
those that form in the parent body; for example, 
\citealt{font,De_Silva_2007,eggen_1970,bubar_king_2010}) and
stellar-orbit diffusion process like radial mixing and radial
migration (for example, \citealt{roskar, quillen}).  After
stars are born---or after a star cluster is accreted and
disrupted---associations or groups will disperse, through
two-body processes, interactions with resonances, or tides from the whole Galaxy.

@@

Six two-dimensional projections of the abundance data, plus some other
data quantities, are shown in \figurename~\ref{fig:all}.  Throughout
this work we show two-dimensional abundance projections for only a few
elements. The entire sequence of 15-choose-2 combinations is too
immense to visualize. The projections shown are ones which
have been extensively used in abundance studies of globular clusters
and satellite systems. These are typically light element abundances,
and our projections include C--N, Na--O, and Mg--Al. Globular clusters
famously demonstrate correlations in these elements \citep[e.g., see][and references therein]{Norris_1995,Carretta_2009}, thereby making
them suitable projections for us to show in verifying and examining
any substructure identified by k-means (or any other clustering algorithm).

@@

All of the code for the data manipulation, clustering, and
visualization is built on \project{numpy} \citep{numpy},
\project{scikit-learn} \citep{sklearn}, and \project{matplotlib}
\citep{matplotlib}.
All of the code written specifically for this project is available in
two open-source code repositories at
\url{https://github.com/andycasey/AnniesLasso} and
\url{https://github.com/davidwhogg/Platypus}.

>>>>

\documentclass[12pt, letterpaper, preprint]{aastex6}

@@

Chemical tagging holds the promise of revealing not just the
star-formation history of the Galaxy, but also the accretion history
(as things that fall in are expected to be chemically distinct from
those that form in the parent body; for example, 
\citealt{eggen_1970,font,De_Silva_2007,bubar_king_2010}) and
stellar-orbit diffusion process like radial mixing and radial
migration (for example, \citealt{roskar, quillen}).  After
stars are born---or after a star cluster is accreted and
disrupted---associations or groups will disperse, through
two-body processes, interactions with resonances, or tides from the whole Galaxy.

@@

Six two-dimensional projections of the abundance data, plus some other
data quantities, are shown in \figurename~\ref{fig:all}.  Throughout
this work we show two-dimensional abundance projections for only a few
elements. The entire sequence of 15-choose-2 combinations is too
immense to visualize. The projections shown are ones which
have been extensively used in abundance studies of globular clusters
and satellite systems. These are typically light element abundances,
and our projections include C--N, Na--O, and Mg--Al. Globular clusters
famously demonstrate correlations in these elements \citep[for example,][and references therein]{Norris_1995,Carretta_2009}, thereby making
them suitable projections for us to show in verifying and examining
any substructure identified by k-means (or any other clustering algorithm).

@@

All of the code written specifically for this project is available in
two open-source code repositories at
\url{https://github.com/andycasey/AnniesLasso} and
\url{https://github.com/davidwhogg/Platypus}.

\software{
	\project{numpy} \citep{numpy},
	\project{scikit-learn}, \citep{sklearn},
	\project{matplotlib} \citep{matplotlib}}.
